### 2015-05-14 whoop.tex
space hack

A \emph{concurrent program} is described by a finite set of \emph{shared} variables, $\sharedvars$, a finite set of mutexes, $\mutexes$, and a finite set of \emph{thread templates}.  A thread template, $T$, consists of a finite set of procedures, $\procedures{T}$, and a finite set of private variables, $\variables{T}$.  A designated procedure, $\mainforthread{T} \in \procedures{T}$, denotes the starting point for execution of $T$ by a thread.  Each procedure of $\procedures{T}$ is represented by a control flow graph of basic blocks, where each block contains a sequence of statements.  A basic block either has a single successor or a pair of successors.  In the latter case, an \emph{exit condition} over thread-private variables determines the successor to which control should flow on block exit.

@@

\noindent\textbf{Semantics }
%
Let $\ids$ be an infinite set from which dynamic thread ids will be drawn.  The state of a running concurrent program consists of: a valuation of the shared variables, $\sharedvars$; a mapping that associates each mutex in $\mutexes$ with an id from $\ids$, recording which thread currently holds the mutex, or with a special value $\bot \notin \ids$ to indicate that the mutex is not held by any thread; and a list of \emph{threads}.  Each thread consists of an id, drawn from $\ids$, a thread template, $T$, an index indicating the next statement of $T$ to be executed by the thread, and a valuation of the thread private variables, $\variables{T}$.  If multiple threads are instances of the same template $T$, then each thread carries a \emph{separate} valuation of the private variables for this template.

@@

\noindent\textbf{Lockset Instrumentation }
%
For thread templates $T$ and $U$ (including the possibility that $T$ and $U$ are equal) we wish to check whether it is possible for a thread executing $T$ to race with a thread executing $U$, in the presence of arbitrarily many further concurrently executing threads.

To achieve this, we first \emph{instrument} a thread template for lockset analysis (see \S\ref{bg:lockset}).  Given an arbitrary symbol, $i$, we define the instrumentation of a template $T$ with respect to $i$, denoted $\instrument{T}{i}$.  There are two aspects to this instrumentation phase: \emph{renaming} and \emph{lockset instrumentation}.

Renaming is straightforward: all occurrences of private variable $x \in \variables{T}$ used in $T$ are replaced with a renamed variable $\instrument{x}{i}$ in $\instrument{T}{i}$, and every procedure $f \in \procedures{T}$ is renamed (both at its declaration site and at all call sites) to $\instrument{f}{i}$ in $\instrument{T}{i}$.  The purpose of renaming is to ensure that when we analyze a pair of templates, $T$ and $U$, both templates execute distinct procedures and operate on distinct private data.  This is vital in the case where $T$ and $U$ are the same.

>>>>

A \emph{concurrent program} is described by a finite set of \emph{shared} variables $\sharedvars$, a finite set of mutexes $\mutexes$, and a finite set of \emph{thread templates}.  A thread template $T$ consists of a finite set of procedures $\procedures{T}$ and a finite set of private variables $\variables{T}$.  A designated procedure $\mainforthread{T} \in \procedures{T}$ denotes the starting point for execution of $T$ by a thread.  Each procedure of $\procedures{T}$ is represented by a control flow graph of basic blocks, where each block contains a sequence of statements.  A basic block either has a single successor or a pair of successors.  In the latter case, an \emph{exit condition} over thread-private variables determines the successor to which control should flow on block exit.

@@

\noindent\textbf{Semantics }
%
Let $\ids$ be an infinite set from which dynamic thread ids will be drawn.  The state of a running concurrent program consists of: a valuation of the shared variables $\sharedvars$; a mapping that associates each mutex in $\mutexes$ with an id from $\ids$, recording which thread currently holds the mutex, or with a special value $\bot \notin \ids$ to indicate that the mutex is not held by any thread; and a list of \emph{threads}.  Each thread consists of an id, drawn from $\ids$, a thread template $T$ an index indicating the next statement of $T$ to be executed by the thread, and a valuation of the thread private variables, $\variables{T}$.  If multiple threads are instances of the same template $T$, then each thread carries a \emph{separate} valuation of the private variables for this template.

@@

\noindent\textbf{Lockset Instrumentation }
%
For templates $T$ and $U$ (including the possibility that $T$ and $U$ are equal) we want to check whether it is possible for a thread executing $T$ to race with a thread executing $U$, in the presence of arbitrarily many further concurrently executing threads.
%
To this end, we first \emph{instrument} a template $T$ for lockset analysis (see \S\ref{bg:lockset}).  Given an arbitrary symbol $i$, we define the instrumentation of $T$ with respect to $i$, denoted $\instrument{T}{i}$.  There are two aspects to this instrumentation phase: \emph{renaming} and \emph{lockset instrumentation}.

Renaming is straightforward: all occurrences of each private variable $x \in \variables{T}$ used in $T$ are replaced with a renamed variable $\instrument{x}{i}$ in $\instrument{T}{i}$, and every procedure $f \in \procedures{T}$ is renamed (both at its declaration site and at all call sites) to $\instrument{f}{i}$ in $\instrument{T}{i}$.  The purpose of renaming is to ensure that when we analyze a pair of templates, $T$ and $U$, both templates execute distinct procedures and operate on distinct private data.  This is vital in the case where $T$ and $U$ are the same.
